---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysis of Video Game Sales
### Jacob Bartlett

## Introduction
Video Games are a popular form of entertainment enjoyed by billions of people worldwide. The games produced range a wide variety of genres and are available to a variety of ages where there's a game for anyone to enjoy. The goal of this analysis is to explore a dataset of video game sales to better understand the consumer market for video games.

In pursuit of the stated goal, I will explore the following questions:

1. Which genres tend to sell more units? Which genres are popular in different regions of the world?

2. How is the ESRB rating related to the number of sales? Does the rating affect the amount of sales?

3. How is the critic rating related to the number of sales? How does the user rating relate?

4. How does the amount of global sales change over the years? Is this affected by the release of new gaming systems/platforms? 

These are the main questions I am looking to answer through the completion of this project. With the findings I'll be able to draw meaningful conclusions on the consumer market of video games throughout the world.

## Data
### Structure
The link to the data set is [here](https://www.kaggle.com/datasets/sidtwr/videogames-sales-dataset/data?select=Video_Games_Sales_as_at_22_Dec_2016.csv). The website contains a csv file of video game sales up until late 2016 with over 16,000 games in the dataset. The number of variables in the dataset is 16. According to the website, the Metacritic scores come from a web scrape of their website. Since Metacritic only reviews games on certain platforms, over half of the games in the dataset are missing values for the critic and user ratings as well as the developer and ESRB rating. 

### Cleaning
I first read in the main csv file I'll be utilizing for this analysis.
```{r, warning = FALSE, message = FALSE}
library(tidyverse)
game_sales <- read_csv('Video_Game_Sales.csv', show_col_types = FALSE)
```

I then started looking at the different values for `Rating` since some of the ESRB rating rules have changed over the years.
```{r, warning = FALSE, message = TRUE}
game_sales %>% group_by(Rating) %>% tally()
```

There are four values that stand out due to having a low number of instances in the dataset. These values are AO, EC, K-A, and RP. AO stands for Adult's Only and is something that is still currently used by gaming companies, but it's very rare for a game to have this rating in general. RP stands for Rating Pending. EC stands for Early Childhood and is meant for ages three and up, but this rating is no longer used and is instead considered E for everyone. K-A stands for Kids to Adults and this rating was changed in 1998 to E for everyone.

To ensure that the variable Rating is properly analyzed I plan on removing the rows with a rating of RP since a pending rating won't provide any valuable insight to an actual rating. The variable EC and K-A will also be changed to E ratings since that's what they would be considered today.
```{r, warning = FALSE, message = FALSE}
game_sales <- game_sales[-c(which(game_sales$Rating == "RP")), ]
game_sales$Rating[game_sales$Rating == "EC"] <- "E"
game_sales$Rating[game_sales$Rating == "K-A"] <- "E"
```

Then I checked the variable types of the dataset to see if any of them needed to be changed to a more suiting type.
```{r, warning = FALSE, message = FALSE}
str(game_sales)
```

Some of the numerical types could be changed to an integer, but that doesn't really effect the data. The only variable whose type needs to be changed is `Year_of_Release` from a character type to an integer or numeric type.
```{r, warning = FALSE, message = FALSE}
game_sales$Year_of_Release <- as.integer(game_sales$Year_of_Release)
```

Finally, in order to make sure I don't run into any issues analyzing the variable `Genre` I checked to see if there were any missing values.
```{r, warning = FALSE, message = FALSE}
which(is.na(game_sales$Genre))
```

There are two missing values for the Genre in row 660 and 14,247, and these same rows are also missing the name of the game. I decided to remove these since there's no way to figure out what game the data is for, and to prevent the missing values from possibly interfering with the analysis of the variable.
```{r, warning = FALSE, message = FALSE}
game_sales <- game_sales[-c(which(is.na(game_sales$Genre))), ]
```

### Variables
* Name: The title of the video game.
* Platform: The gaming platform that the game is on.
* Year_of_Release: The year the video game was released.
* Genre: The main genre of the video game.
* Publisher: The name of the company or corporation that released the game.
* NA_Sales: The total number of physical units sold in millions for North America.
* EU_Sales: The total number of physical units sold in millions for Europe.
* JP_Sales: The total number of physical units sold in millions for Japan.
* Other_Sales: The total number of physical units sold in millions for the rest of the world.
* Global_Sales: The total number of physical units sold in millions.
* Critic_Score: Aggregate score compiled by Metacritic staff.
* Critic_Count: The number of critics used in coming up with the Critic_score.
* User_Score: Score by Metacritic's subscribers.
* User_Count: Number of users who gave the user_score.
* Developer: Party responsible for creating the game.
* Rating: The ESRB rating.

## Results

#### Which genres tend to sell more units? Which genres are popular in different regions of the world?
```{r, warning = FALSE, message = FALSE}
library(patchwork)
p1 <- game_sales %>% group_by(Genre) %>% summarise(total_sales = sum(Global_Sales)) %>% ggplot(aes(x = reorder(Genre, total_sales))) + geom_bar(aes(weight = total_sales)) + coord_flip() + labs(title = "Total Global Sales by Genre", x = "Genre", y = "Total Global Sales (in millions of units)")
p2 <- game_sales %>% group_by(Genre) %>% summarise(Total_NA_Sales = sum(NA_Sales), Total_EU_Sales = sum(EU_Sales), Total_JP_Sales = sum(JP_Sales)) %>% pivot_longer(cols = c(Total_NA_Sales, Total_EU_Sales, Total_JP_Sales), names_to = "Region", values_to = "Sales") %>% ggplot(aes(x = reorder(Genre, Sales), y = Sales)) + geom_bar(stat = "identity") + coord_flip() + facet_grid(. ~ Region) + labs(title = "Total Sales of a Region by Genre", x = "Genre", y = "Total Sales (in millions of units)")
p1 / p2
```

The top graph shows the number of total global sales by genre. The top selling genres for video games are Action, Sports, and Shooter. The series of graphs on the bottom shows which genres are popular within three different regions of the world that have a larger consumer base for video games. Customers living in North America and Europe are more likely to buy games that fit into the Action, Sports, and Shooter genres. The customers living in Japan, however, are more likely to buy games that fit into the Role-Playing, Action, and Sports genre. So, if you're trying to appease more to the Japanese gaming market then making a Role-Playing game would be a more valuable way to go.

#### How is the ESRB rating related to the number of sales? Does the rating affect the amount of sales?
```{r, warning = FALSE, message = FALSE}
game_sales %>% filter(!is.na(Rating), Rating != "AO") %>% group_by(Rating) %>% summarise(average_total_sales = sum(Global_Sales)/n()) %>% ggplot(aes(x = reorder(Rating, average_total_sales), y = average_total_sales)) + geom_bar(stat = "identity") + coord_flip() + labs(titel = "Average Total Global Sales by ESRB Rating", x = "ESRB Rating", y = "Average Total Sales (in millions of units)")
```

The graph above depicts the average total global sales by the ESRB rating of a game. There was an issue with the graph where it shows that games rated AO are more likely to make a lot by over twice as much as any other rating, but there's only one instance of a game with that rating in the dataset so it's not reliable to make that claim. I decided to remove that instance of the dataset to better asses the sales by rating. Games that are rated M for mature are more likely to sell more units on average than any other rating. I predicted that games rated E would sell more on average since they're available to a wider demographic, but given that consumers are more likely to buy action and shooter video games the games that are rated M are more likely to fit into those genres. Games rated E have the second most sales on average though, and this is probably due to the rating of sports video games.

#### How is the critic rating related to the number of sales? How does the user rating relate?
```{r, warning = FALSE, message = FALSE}
p3 <- game_sales %>% filter(!is.na(Critic_Score)) %>% ggplot(aes(x = Critic_Score, y = log(Global_Sales))) + geom_point(alpha = 0.5) + geom_smooth(method = "lm", se = FALSE) + labs(title = "Critic Score by Global Sales", x = "Average Critic Score", y = "Log of Global Sales")
p4 <- game_sales %>% filter(!is.na(User_Score)) %>% ggplot(aes(x = User_Score, y = log(Global_Sales))) + geom_point(alpha = 0.5) + geom_smooth(method = "lm", se = FALSE) + labs(title = "User Score by Global Sales", x = "Average User Score", y = "Log of Global Sales")
p3+p4
```

The graph above on the left depicts the relationship between the average critic score for a game and the number of global sales. The graph on the right depicts the relationship between the average user score for a game and the number of global sales. I predicted that the average user score would have a higher positive correlation to game sales, but the average critic score has the higher positive correlation. So, consumers are more likely to purchase a video game based on the critic ratings rather than the user ratings.

#### How does the amount of global sales change over the years? Is this affected by the release of new gaming systems/platforms? 
```{r, warning = FALSE, message = FALSE}
game_sales %>% filter(!is.na(Year_of_Release), Year_of_Release %in% c(1980:2016)) %>% group_by(Year_of_Release) %>% summarise(Total_Global_Sales = sum(Global_Sales)) %>% ggplot(aes(x = Year_of_Release, y = Total_Global_Sales)) + geom_line() + geom_point() + labs(title = "Total Global Sales by the Year of Release", x = "Year of Release", y = "Total Global Sales (in millions of units)")
```

The graph above depicts the total global sales by the year the game was released. One of the issues with this graph is that the highest year of release is 2016, so the closer a game was released towards the end of 2016 then the lower their sales might be since they haven't had as much time in the market compared to other games. The graph still does show that there was a boom in video game sales starting after 1995. The increase of sales in 1995 can be contributed to the release of the first PlayStation gaming system followed by the release of the Nintendo 64 in 1996. The increase in sales after 2000 can be contributed to the release of the PlayStation 2 in 2000, the Xbox in 2001, and the Nintendo GameCube in 2001. The increase in sales that continues into 2005 and after can be contributed to the release of PlayStation 3 in 2006, the Xbox 360 in 2005, and the Wii in 2006. The subsequent releases of the Xbox One and the PlayStation 4 in the mid 2010's has also boosted gaming sales worldwide, but sadly the dataset is not that recent to show that result

## Conclusion
In conclusion, my analysis on the consumer market for video games has shown several factors that influence the market. One of the main ways to increase the chances of a high selling game is to make a game that appeases to want customers want. As shown in our analysis the top genres that are sold are action, sports, and shooter games. Making games that fit into one, or multiple, of these categories will increase the chances of selling more units for that game. Another factor is the ESRB rating of the game. While games that are rated E can sell well due to being open to a wider age demographic, games that are rated M appeal more to customers since that's what action and shooter game ratings tend to fit into. There's also the importance of how critics react to the game. In order to increase the amount of units sold, companies that allow critics to play their games early in hopes for good reviews will see an increase in the amount of units sold when the game is available on the market. Sometimes the timing of the release also matters. Releasing a game on a new gaming system will also increase sales since newer gaming systems are able to run bigger games with better graphics, and since they're new customers will be looking for new games that are only available for these gaming systems.

Further research on the gaming platform is needed. Since this dataset only contains data as late as December, 2016 it doesn't properly show the popularity of gaming systems like the PlayStation 4 or Xbox One. More recent research would also be useful in general to see if customers are still purchasing the newer gaming systems like the PlayStation 5 and the Xbox Series X, or if they're starting to move more towards video games on the PC. Lastly, there were many variables missing when it came to the critic and user scores as well as the ESRB rating, so it would be interesting if more analysis could be done for comparing the critic and user scores for more video games and from different companies that also have video game reviews. 